---
title: "Thermal Response Functions for Chikungunya Transmission Risk"
author: "Sina Mokhtar"
output:
  pdf_document:
    fig_caption: true
geometry: margin=1in
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Before building the reproduction number model, I first examine the individual temperature-dependent functions used in the model. The purpose is to make sure each function behaves as expected over the temperature range. After checking and approving these functions separately, I will use them in the full $R_0(T)$ model. We use the Mordecai et al. model as the baseline. In this model, the basic reproduction number depends on temperature:

$$
R_0(T)=
\sqrt{
\frac{
a(T)^2 b(T)c(T)e^{-(\frac{\mu(T)}{PDR(T)})}EFD(T)p_{EA}(T)MDR(T)
}{
N r \mu(T)^3
}
}
$$

The temperature-dependent functions are:

$$
a(T) = \text{biting rate}
$$

$$
b(T) = \text{vector-to-human transmission probability}
$$

$$
c(T) = \text{human-to-vector transmission probability}
$$

$$
\mu(T) = \text{adult mosquito mortality rate}
$$

$$
PDR(T) = \text{parasite development rate}
$$

$$
EFD(T) = \text{eggs per female per day}
$$

$$
pEA(T) = \text{egg-to-adult survival probability}
$$

$$
MDR(T) = \text{mosquito development rate}
$$
```{r, echo= F}
library(ggplot2)

# Temperature range
temp <- seq(-10, 50, by = 0.1)
```


\newpage 

# Biting rate

a is the per-mosquito biting rate.

* The Mordecai biting-rate functions were evaluated only where they are positive: $13.35<T<40.08$ for *Ae. aegypti* and $10.25<T<38.32$ for *Ae. albopictus*.
* Ae. albopictus biting rate is $\sqrt{(38.32-T)}$ not $\sqrt{(38.31-T)}$

```{r biting_rate, message=FALSE, warning=FALSE, fig.cap="Biting rate thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Biting Rate ##########


# Species: Ae. aegypti
# Cited as Original Source: Caminade et al., 2017; Focks et al., 1995; Kakarla et al., 2019; Liu-Helmersson et al., 2014; Mclean et al., 1974; Scott et al., 2000
a_aegypti_caminade_2017 <- function(T) {
  
  b_m <- rep(NA_real_, length(T))
  
  idx1 <- T >= 21 & T <= 32
  
  b_m[idx1] <- 0.0943 + 0.0043 * T[idx1]
  
  return(b_m)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
a_aegypti_mordecai_2017 <- function(T) {
  
  a <- rep(NA_real_, length(T))
  
  idx1 <- T > 13.35 & T < 40.08
  
  a[idx1] <- 2.02E-4 * T[idx1] * (T[idx1] - 13.35) * sqrt(40.08 - T[idx1])
  
  return(a)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
a_albopictus_mordecai_2017 <- function(T) {
  
  a <- rep(NA_real_, length(T))
  
  idx1 <- T > 10.25 & T < 38.32
  
  a[idx1] <- 1.93E-4 * T[idx1] * (T[idx1] - 10.25) * sqrt(38.32 - T[idx1])
  
  return(a)
}


# Species: Ae. albopictus
# Cited as Original Source: Caminade et al., 2017; Liu-Helmersson et al., 2014; Scott et al., 2000; Zardini et al., 2024
a_albopictus_caminade_2017 <- function(T) {
  
  a <- 0.5 * (0.0043 * T + 0.0943)
  
  return(a)
}


####### Make data frame ##########

biting_data <- rbind(
  data.frame(
    temp = temp,
    a = a_aegypti_caminade_2017(temp),
    function_name = "Ae. aegypti - Caminade/Focks/..."
  ),
  
  data.frame(
    temp = temp,
    a = a_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    a = a_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    a = a_albopictus_caminade_2017(temp),
    function_name = "Ae. albopictus - Caminade/..."
  )
)



####### Plot biting rate ##########

p_a <- ggplot(biting_data, aes(x = temp, y = a, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 0.4)) +
  labs(
    x = "Temperature (C)",
    y = "Biting rate, a(T)",
    color = "Function",
    title = "Biting rate thermal functions"
  ) +
  theme_bw()

print(p_a)


####### Save figure ##########

ggsave(
  filename = "biting_rate.jpg",
  plot = p_a,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage

# Transmission Probability (Vector -> Human)

b is the proportion of infectious bites that infect susceptible humans.

```{r transmission_vector_to_human, message=FALSE, warning=FALSE, fig.cap="Transmission probability vector to human thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Transmission Probability Vector -> Human ##########


# Species: Ae. aegypti
# Cited as Original Source: Lambrechts et al., 2011
b_aegypti_lambrechts_2011 <- function(T) {
  
  b_prob <- rep(NA_real_, length(T))
  
  idx1 <- T >= 12.4 & T <= 26.1
  idx2 <- T > 26.1 & T <= 32.5
  
  b_prob[idx1] <- -0.9037 + 0.0729 * T[idx1]
  b_prob[idx2] <- 1
  
  return(b_prob)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
b_aegypti_mordecai_2017 <- function(T) {
  
  b_prob <- rep(NA_real_, length(T))
  
  idx1 <- T > 17.05 & T < 35.83
  
  b_prob[idx1] <- 8.49E-4 * T[idx1] * (T[idx1] - 17.05) * sqrt(35.83 - T[idx1])
  
  return(b_prob)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
b_albopictus_mordecai_2017 <- function(T) {
  
  b_prob <- rep(NA_real_, length(T))
  
  idx1 <- T > 15.84 & T < 36.40
  
  b_prob[idx1] <- 7.35E-4 * T[idx1] * (T[idx1] - 15.84) * sqrt(36.40 - T[idx1])
  
  return(b_prob)
}


####### Make data frame ##########

vector_to_human_data <- rbind(
  data.frame(
    temp = temp,
    b_prob = b_aegypti_lambrechts_2011(temp),
    function_name = "Ae. aegypti - Lambrechts 2011"
  ),
  
  data.frame(
    temp = temp,
    b_prob = b_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    b_prob = b_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot transmission probability vector to human ##########

p_b <- ggplot(vector_to_human_data, aes(x = temp, y = b_prob, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Temperature (C)",
    y = "Transmission probability vector to human, b(T)",
    color = "Function",
    title = "Transmission probability vector to human thermal functions"
  ) +
  theme_bw()

print(p_b)


####### Save figure ##########

ggsave(
  filename = "transmission_vector_to_human.jpg",
  plot = p_b,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage


# Transmission Probability (Human -> Vector)

c is the proportion of bites on infected humans that infect previously uninfected mosquitoes. (i.e., b*c = vector competence).

```{r transmission_human_to_vector, message=FALSE, warning=FALSE, fig.cap="Transmission probability human to vector thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Transmission Probability Human -> Vector ##########


# Species: Ae. aegypti
# Cited as Original Source: Lambrechts et al., 2011
c_aegypti_lambrechts_2011 <- function(T) {
  
  c <- rep(NA_real_, length(T))
  
  idx1 <- T > 12.286 & T < 32.461
  
  c[idx1] <- 0.001044 * T[idx1] * (T[idx1] - 12.286) * sqrt(32.461 - T[idx1])
  
  return(c)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
c_aegypti_mordecai_2017 <- function(T) {
  
  c <- rep(NA_real_, length(T))
  
  idx1 <- T > 12.22 & T < 37.46
  
  c[idx1] <- 4.91E-4 * T[idx1] * (T[idx1] - 12.22) * sqrt(37.46 - T[idx1])
  
  return(c)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
c_albopictus_mordecai_2017 <- function(T) {
  
  c <- rep(NA_real_, length(T))
  
  idx1 <- T > 3.62 & T < 36.82
  
  c[idx1] <- 4.39E-4 * T[idx1] * (T[idx1] - 3.62) * sqrt(36.82 - T[idx1])
  
  return(c)
}


####### Make data frame ##########

human_to_vector_data <- rbind(
  data.frame(
    temp = temp,
    c = c_aegypti_lambrechts_2011(temp),
    function_name = "Ae. aegypti - Lambrechts 2011"
  ),
  
  data.frame(
    temp = temp,
    c = c_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    c = c_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot transmission probability human to vector ##########

p_c <- ggplot(human_to_vector_data, aes(x = temp, y = c, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 1.1)) +
  labs(
    x = "Temperature (C)",
    y = "Transmission probability human to vector, c(T)",
    color = "Function",
    title = "Transmission probability human to vector thermal functions"
  ) +
  theme_bw()

print(p_c)


####### Save figure ##########

ggsave(
  filename = "transmission_human_to_vector.jpg",
  plot = p_c,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage

# Adult Mosquito Mortality Rate

$\mu$ is the adult mosquito mortality rate.

For Mordecai et al. 2017, the table gives adult mosquito lifespan, $lf(T)$, not mortality rate directly. Therefore, mortality is calculated as: $\mu(T)=\frac{1}{lf(T)}$

* For Ng et al. 2017, the original source was checked. The $T^5$ coefficient is $-2.72000e^{-8}$, not $-2.72e^{-2}$.

* Mordecai Ae. aegypti functions is $(T-37.73)$ not $(37.73-T)$

* Mordecai albopictus functions is $(T-31.51)$ not $(31.51-T)$
```{r mortality_rate, message=FALSE, warning=FALSE, fig.cap="Adult mosquito mortality rate thermal functions", echo=F, fig.pos="H", out.extra = ""}

####### Adult Mosquito Mortality Rate ##########


# Species: Ae. aegypti
# Cited as Original Source: Caminade et al., 2017
mu_aegypti_caminade_2017 <- function(T) {
  
  mu <- rep(NA_real_, length(T))
  
  idx1 <- T < 22
  idx2 <- T >= 22
  
  mu[idx1] <- 1 / (1.22 + exp(-3.05 + 0.72 * T[idx1])) + 0.196
  mu[idx2] <- 1 / (1.14 + exp(51.4 - 1.3 * T[idx2])) + 0.192
  
  return(mu)
}


# Species: Ae. aegypti
# Cited as Original Source: Focks et al., 1995; Kakarla et al., 2019; Mclean et al., 1974; Yang et al., 2009
mu_aegypti_focks_1995 <- function(T) {
  
  mu <- 0.8692 -
    0.1590 * T +
    0.01116 * T^2 -
    0.0003408 * T^3 +
    0.000003809 * T^4
  
  return(mu)
}


# Species: Ae. albopictus
# Cited as Original Source: Poletti et al., 2011
mu_albopictus_poletti_2011 <- function(T) {
  
  mu <- 0.031 + 95820 * exp(T - 50.4)
  
  return(mu)
}


# Species: Ae. albopictus
# Cited as Original Source: Ruiz-Moreno et al., 2012
mu_albopictus_ruiz_moreno_2012 <- function(T) {
  
  mu <- rep(NA_real_, length(T))
  
  idx1 <- T < -5
  idx2 <- T >= -5 & T < 15
  idx3 <- T >= 15 & T < 38.5
  idx4 <- T >= 38.5
  
  mu[idx1] <- 0
  mu[idx2] <- 1 / (5.6 + 1.12 * T[idx2])
  mu[idx3] <- 1 / (40.27 - 1.38 * T[idx3] + 0.01 * T[idx3]^2)
  mu[idx4] <- 0
  
  return(mu)
}


# Species: Ae. albopictus
# Cited as Original Source: Brady et al., 2013; Caminade et al., 2017; Zardini et al., 2024
mu_albopictus_brady_2013 <- function(T) {
  
  mu <- rep(NA_real_, length(T))
  
  idx1 <- T < 15
  idx2 <- T >= 15 & T < 26.3
  idx3 <- T >= 26.3
  
  mu[idx1] <- 1 / (1.1 + exp(-4.04 + 0.576 * T[idx1])) + 0.12
  mu[idx2] <- 0.000339 * T[idx2]^2 - 0.0189 * T[idx2] + 0.336
  mu[idx3] <- 1 / (1.065 + exp(32.2 - 0.92 * T[idx3])) + 0.0747
  
  return(mu)
}


# Species: Ae. albopictus
# Cited as Original Source: Ng et al., 2017
# Checked Ng et al. (2017): the T^5 coefficient is -2.72000e-08, not -2.72e-02.
mu_albopictus_ng_2017 <- function(T) {
  
  mu <- 1.33048 -
    2.32772e-1 * T +
    1.68529e-2 * T^2 -
    5.61719e-4 * T^3 +
    7.91643e-6 * T^4 -
    2.72e-8 * T^5     # Original Table but shoud modify to 2.72e-8
  
  return(mu)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
# Table gives lifespan; mortality is calculated as mu(T) = 1 / lf(T).
mu_aegypti_mordecai_2017 <- function(T) {
  
  mu <- rep(NA_real_, length(T))
  
  idx1 <- T > 9.16 & T < 37.73
  
  lf <- -1.48e-1 * (T[idx1] - 9.16) * (T[idx1] - 37.73)
  mu[idx1] <- 1 / lf
  
  return(mu)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
# Table gives lifespan; mortality is calculated as mu(T) = 1 / lf(T).
mu_albopictus_mordecai_2017 <- function(T) {
  
  mu <- rep(NA_real_, length(T))
  
  idx1 <- T > 13.41 & T < 31.51
  
  lf <- -1.43 * (T[idx1] - 13.41) * (T[idx1] - 31.51 )
  mu[idx1] <- 1 / lf
  
  return(mu)
}


####### Make data frame ##########

mortality_data <- rbind(
  data.frame(
    temp = temp,
    mu = mu_aegypti_caminade_2017(temp),
    function_name = "Ae. aegypti - Caminade 2017"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_aegypti_focks_1995(temp),
    function_name = "Ae. aegypti - Focks/Kakarla/Mclean/Yang"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_albopictus_poletti_2011(temp),
    function_name = "Ae. albopictus - Poletti 2011"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_albopictus_ruiz_moreno_2012(temp),
    function_name = "Ae. albopictus - Ruiz-Moreno 2012"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_albopictus_brady_2013(temp),
    function_name = "Ae. albopictus - Brady/Caminade/Zardini"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_albopictus_ng_2017(temp),
    function_name = "Ae. albopictus - Ng 2017"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017 from lifespan"
  ),
  
  data.frame(
    temp = temp,
    mu = mu_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017 from lifespan"
  )
)


####### Check possible problems ##########

problem_mu <- mortality_data[
  is.na(mortality_data$mu) |
    !is.finite(mortality_data$mu) |
    mortality_data$mu < 0 |
    mortality_data$mu > 2,
]

problem_summary <- aggregate(
  temp ~ function_name,
  data = problem_mu,
  FUN = length
)

names(problem_summary)[2] <- "number_of_problem_values"


####### Plot mortality rate ##########

p_mu <- ggplot(mortality_data, aes(x = temp, y = mu, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 2)) +
  labs(
    x = "Temperature (C)",
    y = "Adult mosquito mortality rate, mu(T)",
    color = "Function",
    title = "Adult mosquito mortality rate thermal functions"
  ) +
  theme_bw()

print(p_mu)


####### Save figure ##########

ggsave(
  filename = "adult_mosquito_mortality_rate.jpg",
  plot = p_mu,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage


# Parasite Development Rate

PDR is the parasite development rate (i.e., the inverse of the extrinsic incubation period, the time required between a mosquito biting an infected host and becoming infectious)
```{r parasite_development_rate, message=FALSE, warning=FALSE, fig.cap="Parasite development rate thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Extrinsic Incubation Rate / Parasite Development Rate ##########


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
PDR_albopictus_mordecai_2017 <- function(T) {
  
  PDR <- rep(NA_real_, length(T))
  
  idx1 <- T > 10.39 & T < 43.05
  
  PDR[idx1] <- 1.09E-4 * T[idx1] * (T[idx1] - 10.39) * sqrt(43.05 - T[idx1])
  
  return(PDR)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
PDR_aegypti_mordecai_2017 <- function(T) {
  
  PDR <- rep(NA_real_, length(T))
  
  idx1 <- T > 10.68 & T < 45.90
  
  PDR[idx1] <- 6.65E-5 * T[idx1] * (T[idx1] - 10.68) * sqrt(45.90 - T[idx1])
  
  return(PDR)
}


# Species: Ae. aegypti
# Cited as Original Source: Focks et al., 1995
PDR_aegypti_focks_1995 <- function(T) {
  
  PDR <- 1 / (4 + exp(5.15 - 0.123 * T))
  
  return(PDR)
}


# Species: Ae. albopictus
# Cited as Original Source: Chan & Johansson, 2012
PDR_albopictus_chan_johansson_2012 <- function(T) {
  
  PDR <- 1 / exp(log(6) * exp(-0.08 * (T - 28)))
  
  return(PDR)
}


# Species: Ae. albopictus
# Cited as Original Source: Chan & Johansson, 2012
PDR_albopictus_chan_johansson_2012_alt <- function(T) {
  
  PDR <- 1 / (3.0 * exp(-0.21 * (T - 28.0)))
  
  return(PDR)
}


# Species: Ae. albopictus
# Cited as Original Source: Ruiz-Moreno et al., 2012
PDR_albopictus_ruiz_moreno_2012 <- function(T) {
  
  EIP <- rep(NA_real_, length(T))
  
  idx1 <- T < 10
  idx2 <- T >= 10 & T <= 32
  idx3 <- T > 32
  
  EIP[idx1] <- 4
  EIP[idx2] <- 113 / 22 - (2.4 / 22) * T[idx2]
  EIP[idx3] <- 1.5
  
  PDR <- 1 / EIP
  
  return(PDR)
}


####### Make data frame ##########

PDR_data <- rbind(
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_aegypti_focks_1995(temp),
    function_name = "Ae. aegypti - Focks 1995"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_chan_johansson_2012(temp),
    function_name = "Ae. albopictus - Chan 2012"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_chan_johansson_2012_alt(temp),
    function_name = "Ae. albopictus - Chan 2012 alt"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_ruiz_moreno_2012(temp),
    function_name = "Ae. albopictus - Ruiz 2012"
  )
)


####### Plot parasite development rate ##########

p_PDR <- ggplot(PDR_data, aes(x = temp, y = PDR, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 0.3)) +
  labs(
    x = "Temperature (C)",
    y = "Parasite development rate, PDR(T)",
    color = "Function",
    title = "Parasite development rate thermal functions"
  ) +
  theme_bw()

print(p_PDR)


####### Save figure ##########

ggsave(
  filename = "parasite_development_rate.jpg",
  plot = p_PDR,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```

\newpage

# Eggs per Female per Day (Fecundity)

EFD (Ae. aegypti) is the number of eggs produced per female mosquito per day and TFD (Ae. albopictus) is number of eggs laid per female per
gonotrophic cycle (number/female)



```{r fecundity, message=FALSE, warning=FALSE, fig.cap="Fecundity thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Fecundity ##########


# Species: Ae. aegypti
# Cited as Original Source: Yang et al., 2009
EFD_aegypti_yang_2009 <- function(T) {
  
  EFD <- -341.8585 +
    108.7373 * T -
    12.3447 * T^2 +
    0.6367451 * T^3 -
    0.0149469 * T^4 +
    0.0001294166 * T^5
  
  return(EFD)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
EFD_albopictus_mordecai_2017 <- function(T) {
  
  EFD <- rep(NA_real_, length(T))
  
  idx1 <- T > 8.02 & T < 35.65
  
  EFD[idx1] <- 4.88E-02 * T[idx1] * (T[idx1] - 8.02) * sqrt(35.65 - T[idx1])
  
  return(EFD)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
EFD_aegypti_mordecai_2017 <- function(T) {
  
  EFD <- rep(NA_real_, length(T))
  
  idx1 <- T > 14.58 & T < 34.61
  
  EFD[idx1] <- 8.56E-03 * T[idx1] * (T[idx1] - 14.58) * sqrt(34.61 - T[idx1])
  
  return(EFD)
}


####### Make data frame ##########

fecundity_data <- rbind(
  data.frame(
    temp = temp,
    EFD = EFD_aegypti_yang_2009(temp),
    function_name = "Ae. aegypti - Yang 2009"
  ),
  
  data.frame(
    temp = temp,
    EFD = EFD_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    EFD = EFD_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  )
)


####### Plot fecundity ##########

p_EFD <- ggplot(fecundity_data, aes(x = temp, y = EFD, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Fecundity, EFD(T)",
    color = "Function",
    title = "Fecundity thermal functions"
  ) +
  theme_bw()

print(p_EFD)


####### Save figure ##########

ggsave(
  filename = "fecundity.jpg",
  plot = p_EFD,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```




\newpage 

# Egg-Adult Survival Probability

$p_{EA}$ is the mosquito egg-to-adult survival probability

* The *Ae. albopictus* Mordecai function is $(T - 39.33)$ not $(39.33-T)$.

```{r egg_adult_survival_probability, message=FALSE, warning=FALSE, fig.cap="Egg-adult survival probability thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Egg-Adult Survival Probability ##########


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
pEA_aegypti_mordecai_2017 <- function(T) {
  
  pEA <- rep(NA_real_, length(T))
  
  idx1 <- T > 13.56 & T < 38.29
  
  pEA[idx1] <- 5.99E-03 * (T[idx1] - 13.56) * (38.29 - T[idx1])
  
  return(pEA)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
pEA_albopictus_mordecai_2017 <- function(T) {
  
  pEA <- rep(NA_real_, length(T))
  
  idx1 <- T > 9.04 & T < 39.33
  
  pEA[idx1] <- -3.61E-03 * (T[idx1] - 9.04) * (T[idx1] - 39.33)
  
  return(pEA)
}


####### Make data frame ##########

pEA_data <- rbind(
  data.frame(
    temp = temp,
    pEA = pEA_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    pEA = pEA_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot egg-adult survival probability ##########

p_pEA <- ggplot(pEA_data, aes(x = temp, y = pEA, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Egg-adult survival probability, pEA(T)",
    color = "Function",
    title = "Egg-adult survival probability thermal functions"
  ) +
  theme_bw()

print(p_pEA)


####### Save figure ##########

ggsave(
  filename = "egg_adult_survival_probability.jpg",
  plot = p_pEA,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```

\newpage

# Mosquito Immature Development Rate (Aquatic Maturation)

MDR is the mosquito immature development rate (i.e., the inverse of the egg-to-adult development time)

```{r aquatic_maturation, message=FALSE, warning=FALSE, fig.cap="Aquatic maturation thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Aquatic Maturation ##########


# Species: Ae. aegypti
# Cited as Original Source: Zhu et al., 2019
phi_v_aegypti_zhu_2019 <- function(T) {
  
  phi_v <- (
    0.131 -
      0.05723 * T +
      0.01164 * T^2 -
      0.001341 * T^3 +
      8.723E-5 * T^4 -
      3.017E-6 * T^5 +
      5.153E-8 * T^6 -
      3.42E-10 * T^7
  ) / 7
  
  return(phi_v)
}


# Species: Ae. aegypti
# Cited as Original Source: Mordecai et al., 2017
MDR_aegypti_mordecai_2017 <- function(T) {
  
  MDR <- rep(NA_real_, length(T))
  
  idx1 <- T > 11.36 & T < 39.17
  
  MDR[idx1] <- 7.86E-5 * T[idx1] * (T[idx1] - 11.36) * sqrt(39.17 - T[idx1])
  
  return(MDR)
}


# Species: Ae. albopictus
# Cited as Original Source: Mordecai et al., 2017
MDR_albopictus_mordecai_2017 <- function(T) {
  
  MDR <- rep(NA_real_, length(T))
  
  idx1 <- T > 8.60 & T < 39.66
  
  MDR[idx1] <- 6.38E-5 * T[idx1] * (T[idx1] - 8.60) * sqrt(39.66 - T[idx1])
  
  return(MDR)
}


####### Make data frame ##########

aquatic_maturation_data <- rbind(
  data.frame(
    temp = temp,
    MDR = phi_v_aegypti_zhu_2019(temp),
    function_name = "Ae. aegypti - Zhu 2019"
  ),
  
  data.frame(
    temp = temp,
    MDR = MDR_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    MDR = MDR_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot aquatic maturation ##########

p_MDR <- ggplot(aquatic_maturation_data, aes(x = temp, y = MDR, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Aquatic maturation, MDR(T)",
    color = "Function",
    title = "Aquatic maturation thermal functions"
  ) +
  theme_bw()

print(p_MDR)


####### Save figure ##########

ggsave(
  filename = "aquatic_maturation.jpg",
  plot = p_MDR,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```





