---
title: "Thermal Response Functions for Chikungunya Transmission Risk"
author: "Sina Mokhtar"
output:
  pdf_document:
    fig_caption: true
geometry: margin=1in
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Before building the reproduction number model, I first examine the individual temperature-dependent functions used in the model. The purpose is to make sure each function behaves as expected over the temperature range. After checking and approving these functions separately, I will use them in the full $R_0(T)$ model. We use the Mordecai et al. model as the baseline. In this model, the basic reproduction number depends on temperature:

$$
R_0(T)=
\sqrt{
\frac{
a(T)^2 b(T)c(T)e^{-(\frac{\mu(T)}{PDR(T)})}EFD(T)p_{EA}(T)MDR(T)
}{
N r \mu(T)^3
}
}
$$

The temperature-dependent functions are:

$$
a(T) = \text{biting rate}
$$

$$
b(T) = \text{vector-to-human transmission probability}
$$

$$
c(T) = \text{human-to-vector transmission probability}
$$

$$
\mu(T) = \text{adult mosquito mortality rate}
$$

$$
PDR(T) = \text{parasite development rate}
$$

$$
EFD(T) = \text{eggs per female per day}
$$

$$
pEA(T) = \text{egg-to-adult survival probability}
$$

$$
MDR(T) = \text{mosquito development rate}
$$
```{r, echo= F}
library(ggplot2)
source("../R/thermal_functions.R")

# Temperature range
temp <- seq(15, 36, by = 0.1)

# Output directories are relative to this Rmd in scripts/.
fig_dir <- "../figures"
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)
```


\newpage 

# Biting rate

a is the per-mosquito biting rate.

* For the Scott et al. (2000) *Ae. aegypti* function, values outside 21--32 C are set to zero.

```{r biting_rate, message=FALSE, warning=FALSE, fig.cap="Biting rate thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Biting Rate ##########


# Functions are sourced from ../R/thermal_functions.R


####### Make data frame ##########

biting_data <- rbind(
  data.frame(
    temp = temp,
    a = a_aegypti_scott_2000(temp),
    function_name = "Ae. aegypti - Scott 2000"
  ),
  
  data.frame(
    temp = temp,
    a = a_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    a = a_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    a = a_albopictus_scott_2000(temp),
    function_name = "Ae. albopictus - Scott 2000"
  )
)



####### Plot biting rate ##########

p_a <- ggplot(biting_data, aes(x = temp, y = a, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 0.4)) +
  labs(
    x = "Temperature (C)",
    y = "Biting rate, a(T)",
    color = "Function",
    title = "Biting rate thermal functions"
  ) +
  theme_bw()

print(p_a)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "biting_rate.jpg"),
  plot = p_a,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage

# Transmission Probability (Vector -> Human)

b is the proportion of infectious bites that infect susceptible humans.

```{r transmission_vector_to_human, message=FALSE, warning=FALSE, fig.cap="Transmission probability vector to human thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Transmission Probability Vector -> Human ##########


# Functions are sourced from ../R/thermal_functions.R


####### Make data frame ##########

vector_to_human_data <- rbind(
  data.frame(
    temp = temp,
    b_prob = b_aegypti_lambrechts_2011(temp),
    function_name = "Ae. aegypti - Lambrechts 2011"
  ),
  
  data.frame(
    temp = temp,
    b_prob = b_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    b_prob = b_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot transmission probability vector to human ##########

p_b <- ggplot(vector_to_human_data, aes(x = temp, y = b_prob, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Temperature (C)",
    y = "Transmission probability vector to human, b(T)",
    color = "Function",
    title = "Transmission probability vector to human thermal functions"
  ) +
  theme_bw()

print(p_b)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "transmission_vector_to_human.jpg"),
  plot = p_b,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage


# Transmission Probability (Human -> Vector)

c is the proportion of bites on infected humans that infect previously uninfected mosquitoes. (i.e., b*c = vector competence).

```{r transmission_human_to_vector, message=FALSE, warning=FALSE, fig.cap="Transmission probability human to vector thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Transmission Probability Human -> Vector ##########


# Functions are sourced from ../R/thermal_functions.R


####### Make data frame ##########

human_to_vector_data <- rbind(
  data.frame(
    temp = temp,
    c = c_aegypti_lambrechts_2011(temp),
    function_name = "Ae. aegypti - Lambrechts 2011"
  ),
  
  data.frame(
    temp = temp,
    c = c_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    c = c_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot transmission probability human to vector ##########

p_c <- ggplot(human_to_vector_data, aes(x = temp, y = c, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 1.1)) +
  labs(
    x = "Temperature (C)",
    y = "Transmission probability human to vector, c(T)",
    color = "Function",
    title = "Transmission probability human to vector thermal functions"
  ) +
  theme_bw()

print(p_c)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "transmission_human_to_vector.jpg"),
  plot = p_c,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage

# Adult Mosquito Lifespan

$1/\mu(T)$ is the adult mosquito lifespan.

* The correct *Ae. albopictus* lifespan formula is $-1.43(T-13.41)(T-31.51)$, not $-1.43T(T-13.41)(T-31.51)$.

```{r lifespan, message=FALSE, warning=FALSE, fig.cap="Adult mosquito lifespan thermal functions", echo=F, fig.pos="H", out.extra = ""}

####### Adult Mosquito Lifespan ##########



# Functions are sourced from ../R/thermal_functions.R



####### Make data frame ##########

lifespan_data <- rbind(
  data.frame(
    temp = temp,
    lf = lf_aegypti_brady_2013(temp),
    function_name = "Ae. aegypti - Brady 2013"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_aegypti_yang_2009(temp),
    function_name = "Ae. aegypti - Yang 2009"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_albopictus_poletti_2011(temp),
    function_name = "Ae. albopictus - Poletti 2011"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_albopictus_ruiz_moreno_2012(temp),
    function_name = "Ae. albopictus - Ruiz-Moreno 2012"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_albopictus_brady_2013(temp),
    function_name = "Ae. albopictus - Brady 2013"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_albopictus_ng_2017(temp),
    function_name = "Ae. albopictus - Ng 2017"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    lf = lf_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  )
)


####### Plot lifespan ##########

p_lf <- ggplot(lifespan_data, aes(x = temp, y = lf, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Adult mosquito lifespan, 1/mu(T)",
    color = "Function",
    title = "Adult mosquito lifespan thermal functions"
  ) +
  theme_bw()

print(p_lf)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "adult_mosquito_lifespan.jpg"),
  plot = p_lf,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage


# Parasite Development Rate

PDR is the parasite development rate (i.e., the inverse of the extrinsic incubation period, the time required between a mosquito biting an infected host and becoming infectious)
```{r parasite_development_rate, message=FALSE, warning=FALSE, fig.cap="Parasite development rate thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Extrinsic Incubation Rate / Parasite Development Rate ##########



# Functions are sourced from ../R/thermal_functions.R



####### Make data frame ##########

PDR_data <- rbind(
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_aegypti_focks_1995(temp),
    function_name = "Ae. aegypti - Focks 1995"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_chan_johansson_2012(temp),
    function_name = "Ae. albopictus - Chan & Johansson 2012"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_chan_johansson_2012_alt(temp),
    function_name = "Ae. albopictus - Chan & Johansson 2012 alt"
  ),
  
  data.frame(
    temp = temp,
    PDR = PDR_albopictus_ruiz_moreno_2012(temp),
    function_name = "Ae. albopictus - Ruiz-Moreno 2012"
  )
)


####### Plot parasite development rate ##########

p_PDR <- ggplot(PDR_data, aes(x = temp, y = PDR, color = function_name)) +
  geom_line(linewidth = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Temperature (C)",
    y = "Parasite development rate, PDR(T)",
    color = "Function",
    title = "Parasite development rate thermal functions"
  ) +
  theme_bw()

print(p_PDR)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "parasite_development_rate.jpg"),
  plot = p_PDR,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```

\newpage

# Eggs per Female per Day (Fecundity)

EFD (Ae. aegypti) is the number of eggs produced per female mosquito per day and TFD (Ae. albopictus) is number of eggs laid per female per
gonotrophic cycle (number/female)



```{r fecundity, message=FALSE, warning=FALSE, fig.cap="Fecundity thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Fecundity ##########



# Functions are sourced from ../R/thermal_functions.R



####### Make data frame ##########

fecundity_data <- rbind(
  data.frame(
    temp = temp,
    EFD = EFD_aegypti_yang_2009(temp),
    function_name = "Ae. aegypti - Yang 2009"
  ),
  
  data.frame(
    temp = temp,
    EFD = TFD_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    EFD = EFD_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  )
)


####### Plot fecundity ##########

p_EFD <- ggplot(fecundity_data, aes(x = temp, y = EFD, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Fecundity",
    color = "Function",
    title = "Fecundity thermal functions"
  ) +
  theme_bw()

print(p_EFD)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "fecundity.jpg"),
  plot = p_EFD,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```




\newpage 

# Egg-Adult Survival Probability

$p_{EA}$ is the mosquito egg-to-adult survival probability


```{r egg_adult_survival_probability, message=FALSE, warning=FALSE, fig.cap="Egg-adult survival probability thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Egg-Adult Survival Probability ##########



# Functions are sourced from ../R/thermal_functions.R



####### Make data frame ##########

pEA_data <- rbind(
  data.frame(
    temp = temp,
    pEA = pEA_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    pEA = pEA_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    pEA = pEA_albopictus_yang_2009(temp),
    function_name = "Ae. albopictus - Yang 2009"
  ),
  
  data.frame(
    temp = temp,
    pEA = pEA_albopictus_poletti_2011(temp),
    function_name = "Ae. albopictus - Poletti 2011"
  )
)


####### Plot egg-adult survival probability ##########

p_pEA <- ggplot(pEA_data, aes(x = temp, y = pEA, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Egg-adult survival probability, pEA(T)",
    color = "Function",
    title = "Egg-adult survival probability thermal functions"
  ) +
  theme_bw()

print(p_pEA)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "egg_adult_survival_probability.jpg"),
  plot = p_pEA,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```

\newpage

# Mosquito Immature Development Rate (Aquatic Maturation)

MDR is the mosquito immature development rate (i.e., the inverse of the egg-to-adult development time)

```{r aquatic_maturation, message=FALSE, warning=FALSE, fig.cap="Aquatic maturation thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Aquatic Maturation ##########



# Functions are sourced from ../R/thermal_functions.R



####### Make data frame ##########

aquatic_maturation_data <- rbind(
  data.frame(
    temp = temp,
    MDR = phi_v_aegypti_yang_2009(temp),
    function_name = "Ae. aegypti - Yang 2009"
  ),
  
  data.frame(
    temp = temp,
    MDR = MDR_aegypti_mordecai_2017(temp),
    function_name = "Ae. aegypti - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    MDR = MDR_albopictus_mordecai_2017(temp),
    function_name = "Ae. albopictus - Mordecai 2017"
  ),
  
  data.frame(
    temp = temp,
    MDR = MDR_albopictus_poletti_2011(temp),
    function_name = "Ae. albopictus - Poletti 2011"
  )
)


####### Plot aquatic maturation ##########

p_MDR <- ggplot(aquatic_maturation_data, aes(x = temp, y = MDR, color = function_name)) +
  geom_line(linewidth = 1) +
  labs(
    x = "Temperature (C)",
    y = "Aquatic maturation, MDR(T)",
    color = "Function",
    title = "Aquatic maturation thermal functions"
  ) +
  theme_bw()

print(p_MDR)


####### Save figure ##########

ggsave(
  filename = file.path(fig_dir, "aquatic_maturation.jpg"),
  plot = p_MDR,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```


\newpage

# Mordecai R0

This section combines the default Mordecai et al. 2017 thermal functions into the baseline temperature-dependent $R_0(T)$ expression.



```{r mordecai_r0, message=FALSE, warning=FALSE, fig.cap="Mordecai R0 thermal functions", echo=F, fig.pos="H", out.extra = ""}
####### Mordecai R0 ##########



# R0 function is sourced from ../R/thermal_functions.R



####### Calculate R0 ##########

r0_aegypti_data <- data.frame(
  temp = temp,
  R0 = relative_R0_mordecai(
    a = a_aegypti_mordecai_2017(temp),
    b = b_aegypti_mordecai_2017(temp),
    c = c_aegypti_mordecai_2017(temp),
    mu = 1 / lf_aegypti_mordecai_2017(temp),
    PDR = PDR_aegypti_mordecai_2017(temp),
    EFD = EFD_aegypti_mordecai_2017(temp),
    pEA = pEA_aegypti_mordecai_2017(temp),
    MDR = MDR_aegypti_mordecai_2017(temp)
  )
)

r0_aegypti_data$R0[!is.finite(r0_aegypti_data$R0)] <- NA
r0_aegypti_data$R0 <- r0_aegypti_data$R0 / max(r0_aegypti_data$R0, na.rm = TRUE)

r0_albopictus_data <- data.frame(
  temp = temp,
  R0 = relative_R0_mordecai(
    a = a_albopictus_mordecai_2017(temp),
    b = b_albopictus_mordecai_2017(temp),
    c = c_albopictus_mordecai_2017(temp),
    mu = 1 / lf_albopictus_mordecai_2017(temp),
    PDR = PDR_albopictus_mordecai_2017(temp),
    EFD = TFD_albopictus_mordecai_2017(temp),
    pEA = pEA_albopictus_mordecai_2017(temp),
    MDR = MDR_albopictus_mordecai_2017(temp)
  )
)

r0_albopictus_data$R0[!is.finite(r0_albopictus_data$R0)] <- NA
r0_albopictus_data$R0 <- r0_albopictus_data$R0 / max(r0_albopictus_data$R0, na.rm = TRUE)


####### Plot Ae. aegypti R0 ##########

p_r0_aegypti <- ggplot(r0_aegypti_data, aes(x = temp, y = R0)) +
  geom_line(linewidth = 1, color = "#1f78b4") +
  labs(
    x = "Temperature (C)",
    y = "Standardized relative R0(T)",
    title = "Standardized Mordecai R0 thermal function for Ae. aegypti"
  ) +
  theme_bw()

print(p_r0_aegypti)


####### Save Ae. aegypti R0 figure ##########

ggsave(
  filename = file.path(fig_dir, "r0_aegypti_mordecai.jpg"),
  plot = p_r0_aegypti,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)


####### Plot Ae. albopictus R0 ##########

p_r0_albopictus <- ggplot(r0_albopictus_data, aes(x = temp, y = R0)) +
  geom_line(linewidth = 1, color = "#33a02c") +
  labs(
    x = "Temperature (C)",
    y = "Standardized relative R0(T)",
    title = "Standardized Mordecai R0 thermal function for Ae. albopictus"
  ) +
  theme_bw()

print(p_r0_albopictus)


####### Save Ae. albopictus R0 figure ##########

ggsave(
  filename = file.path(fig_dir, "r0_albopictus_mordecai.jpg"),
  plot = p_r0_albopictus,
  width = 8,
  height = 5,
  dpi = 300,
  device = "jpeg",
  bg = "white"
)
```





